Bird beak's elaboration and innovation

### 2022-02-21 Analysis/10-Figure1.Rmd
Update 10-Figure1.Rmd

    ## Get the correct group
    if(length(ID <- which(names(test) %in% names(name.col))) < 1) {
        ID <- which(names(test) %in% paste0(names(name.col), ":phylogeny"))
    }
    elaboration <- c(elaboration[[ID]])
    exploration <- c(exploration[[ID]])

@@

```{r}
box_size <- 1.5
pdf(file = "../Manuscript/Figures/Figure_1b.pdf", height = 2*5, width = box_size*12)
## Global window

layout_matrix <- layout(cbind(
                        rbind(matrix(1, ncol = 2, nrow = 3),
                              matrix(2:5, ncol = 2, nrow = 2, byrow = TRUE)),
                        matrix(6:((27*2)+1), ncol = 10, nrow = 5, byrow = TRUE)
                        ))

@@

## Plotting the tree

tree <- ladderize(tree)
## Get the tip_values
orders_lvl <- shapespace$level2
superorders_lvl <- shapespace$level1

@@

## Setting the tips edges colours
tip_edges <- match.tip.edge(tree_short$tip.label, tree_short, replace.na = "grey")
## Rainbow the tip edges
tip_edges[!(tip_edges %in% "grey")] <- rev(colour.palette(Ntip(tree_short)))
## Plotting the tree
par(mar = c(0, 0, 0, 0)+0.1)
plot(tree_short, show.tip.label = TRUE, edge.color = tip_edges, edge.width = 3, cex = 1)
# tiplabels()

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Tinamiformes"= 1,
    "Galliformes"= 3,
    "Anseriformes"= 2,
    "Otidiformes"= 7,
    "Cuculiformes" = 8,
    "Gruiformes" = 9, 
    "Musophagiformes" = 10,
    "Procellariiformes" = 12,
    "Sphenisciformes" = 11,
    "Ciconiiformes" = 13,
    "Pelecaniformes"= 15,
    "Suliformes" = 14,
    "Podicipediformes" = 4,
    "Columbiformes" = 6,
    "Pterocliformes" = 5,
    "Charadriiformes" = 18,
    "Trogoniformes" = 21,
    "Bucerotiformes" = 22,
    "Coraciiformes" = 23,
    "Piciformes" = 24,
    "Strigiformes" = 20,
    "Accipitriformes" = 19, 
    "Falconiformes" = 25,
    "Psittaciformes" = 26,
    "Passeriformes" = 27,
    "Caprimulgiformes" = 16, 
    "Apodiformes" = 17)
tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))
# par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
# plot(rep(1,27), 1:27, col = "white")
# labs <- names(tip_colours)
# cols <- tip_colours
# text(x = rep(1,27), y = 1:27, label = rev(labs), col = rev(cols))

>>>>

    ## Get the correct group
    if(length(ID <- which(names(elaboration) %in% names(name.col))) < 1) {
        ID <- which(names(elaboration) %in% paste0(names(name.col), ":phylogeny"))
    }
    elaboration <- c(elaboration[[ID]])
    exploration <- c(exploration[[ID]])

@@

```{r}
box_size <- 1.5
pdf(file = "../Manuscript/Figures/Figure_1.pdf", height = box_size*8, width = box_size*10)

## Global window
lay_mat_order <- rbind(
        matrix(c(1,1, 2:9), ncol = 10, nrow = 1), # line 1
        matrix(c(1,1, 10:17), ncol = 10, nrow = 1), # line 2
        matrix(c(1,1, 18:25), ncol = 10, nrow = 1), # line 3
        matrix(c(26:35), ncol = 10, nrow = 1), # line 4
        matrix(c(36:45), ncol = 10, nrow = 1), # line 5
        matrix(c(46:55), ncol = 10, nrow = 1) # line 7
    )
lay_mat_super <- rbind(
        matrix(c(1,1, 2:9), ncol = 10, nrow = 1), # line 7
        matrix(c(1,1, 10:17), ncol = 10, nrow = 1) # line 8
    )

layout_matrix <- layout(rbind(lay_mat_super, (lay_mat_order+17)))

@@

## PLOTTING THE SUPER ORDERS
tree <- ladderize(tree)
## Get the tip_values
orders_lvl <- shapespace$level2
superorders_lvl <- shapespace$level1

## Select the level
selected_level <- coloured_levels <- superorders_lvl

@@

## Setting the tips edges colours
tip_edges <- match.tip.edge(tree_short$tip.label, tree_short, replace.na = "grey")
## Rainbow the tip edges
tip_edges[!(tip_edges %in% "grey")] <- rev(colour.palette(Ntip(tree_short)))
## Adding the number of tips
tree_plot <- tree_short
n_species <- table(selected_level)
for(i in 1:Ntip(tree_plot)) {
    ## Get the number of species
    n_sp <- n_species[which(names(n_species) %in% tree_plot$tip.label[i])]
    tree_plot$tip.label[i] <- paste0(tree_plot$tip.label[i], " (", n_sp, " sp)")
}

## Plotting the tree
par(mar = c(0, 0, 0, 0)+0.1)
plot(tree_plot, show.tip.label = TRUE, edge.color = tip_edges, edge.width = 3, cex = 1)
# tiplabels()

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Paleognathae" = 1,
    "Galloanserae" = 2,
    "Otidimorphae" = 5,
    "Aequornithes" = 6,
    "Mirandornithes" = 3,
    "Columbimorphae" = 4,
    "Telluraves" = 8,
    "Strisores" = 7)
tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))
# par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
# plot(rep(1,27), 1:27, col = "white")
# labs <- names(tip_colours)
# cols <- tip_colours
# text(x = rep(1,27), y = 1:27, label = rev(labs), col = rev(cols))

    ## Plot the ellipse
    par(bty = "n", mar = c(1, 1, 1, 1)+0.1)
    plot.one.ellipse(superorder_shapespace, name.col, x = FALSE, y = FALSE, with.phylo = TRUE)
    text(0,1.2, names(name.col), cex = 1.5, col = name.col)

    ## Plot the elaboration/exploration
    par(bty = "n", mar = c(2, 1, 1, 1)+0.1)
    plot.one.proj.rej(results$super_results_between, name.col, x = TRUE, y = FALSE)
}

## PLOTTING THE ORDERS
tree <- ladderize(tree)
## Get the tip_values
orders_lvl <- shapespace$level2
superorders_lvl <- shapespace$level1

## Replace the order levels as colours
colour_palette <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))
levels(coloured_levels) <- c("grey", colour.palette(length(levels(coloured_levels)[-1])))

## Setting the tips edges colours
tip_edges <- match.tip.edge(tree_short$tip.label, tree_short, replace.na = "grey")
## Rainbow the tip edges
tip_edges[!(tip_edges %in% "grey")] <- rev(colour.palette(Ntip(tree_short)))

## Adding the number of tips
tree_plot <- tree_short
n_species <- table(selected_level)
for(i in 1:Ntip(tree_plot)) {
    ## Get the number of species
    n_sp <- n_species[which(names(n_species) %in% tree_plot$tip.label[i])]
    tree_plot$tip.label[i] <- paste0(tree_plot$tip.label[i], " (", n_sp, " sp)")
}

## Plotting the tree
par(mar = c(0, 0, 0, 0)+0.1)
plot(tree_plot, show.tip.label = TRUE, edge.color = tip_edges, edge.width = 3, cex = 1)

## Matching the order and the colours to the ladderised tree
tip_order <- c(
    "Tinamiformes"= 1,
    "Galliformes"= 3,
    "Anseriformes"= 2,
    "Otidiformes"= 7,
    "Cuculiformes" = 8,
    "Gruiformes" = 9, 
    "Musophagiformes" = 10,
    "Procellariiformes" = 12,
    "Sphenisciformes" = 11,
    "Ciconiiformes" = 13,
    "Pelecaniformes"= 15,
    "Suliformes" = 14,
    "Podicipediformes" = 4,
    "Columbiformes" = 6,
    "Pterocliformes" = 5,
    "Charadriiformes" = 18,
    "Trogoniformes" = 21,
    "Bucerotiformes" = 22,
    "Coraciiformes" = 23,
    "Piciformes" = 24,
    "Strigiformes" = 20,
    "Accipitriformes" = 19, 
    "Falconiformes" = 25,
    "Psittaciformes" = 26,
    "Passeriformes" = 27,
    "Caprimulgiformes" = 16, 
    "Apodiformes" = 17)
tip_colours <- colour.palette(Ntip(tree_short))
names(tip_colours) <- names(sort(tip_order))
